### 2018-06-17 express2018/conclusions.tex
text

We have highlighted a formalisation of the theory of CCS in the 
HOL4 theorem prover (for lack of space we have not discussed 
the formalisation of some basic algebraic theory, of the basic
properties of the expansion preorder,   and of a few
 versions of `bisimulation up to'
techniques). % such as  bisimulation up-to bisimilarity). 
The formalisation has focused on the theory of
unique solution of equations and contractions. 
It has also allowed us to further develop the theory,
notably the basic properties of rooted contraction, and the unique
solution theorem for it with respect to rooted bisimilarity. 
The formalisation brings up and exploits similarities between results
and proofs for different equivalences and preorders. 
We think that the statements in the formalisation are easy to read and
understand, as they remain very close  to  the ordinary  statements of
the theory of CCS. 
%Moreover, 
%the logic foundations of HOL makes the whole work (or individual
%parts) easily portable to other theorem provers.

As future work,  it would be worth extending the current work
to multiple equations and contractions; and considering other equivalences
and preorders, notably trace relations.  
%On another line, one could  
%examine the formalisation of  a different approach to unique
%solutions, by Durier et al.\ \cite{DurierHS17}, in which the use of contraction is
%replaced by semantic conditions on process transitions such as
%divergence.

@@

We would like to thank many people from the HOL community:
Michael Norrish, Thomas Tuerk, 
Ramana Kumar, Konrad Slind, Jeremy Dawson, Robert Beers, etc.
% The second half of this
% paper was written in memory of  Mike J. Gordon.

>>>>

We have highlighted a formalisation of the theory of CCS in the 
HOL4 theorem prover (for lack of space we have not discussed 
the formalisation of some basic algebraic theory, of the basic
properties of the expansion preorder,   and of a few
 versions of `bisimulation up to'
techniques). % such as  bisimulation up-to bisimilarity). 
The formalisation has focused on the theory of
unique solution of equations and contractions. 
It has also allowed us to further develop the theory,
notably the basic properties of rooted contraction, and the unique
solution theorem for it with respect to rooted bisimilarity. 
The formalisation brings up and exploits similarities between results
and proofs for different equivalences and preorders. 
We think that the statements in the formalisation are easy to read and
understand, as they remain very close  to  the ordinary  statements of
the theory of CCS. 
Moreover, 
the logic foundations of HOL makes the whole work (or individual
parts) easily portable to other theorem provers.

As future work,  it would be worth extending the current work
to multiple equations and contractions; and considering other equivalences
and preorders, notably trace relations.  
On another line, one could  
examine the formalisation of  a different approach to unique
solutions, by Durier et al.\ \cite{DurierHS17}, in which the use of contraction is
replaced by semantic conditions on process transitions such as
divergence.

@@

We have benefitted from suggestions and comments 
from several people from the HOL
community, including Robert Beers, Jeremy Dawson,
Ramana Kumar,
Michael Norrish, 
 Konrad Slind,
Thomas Tuerk.

% The second half of this
% paper was written in memory of  Mike J. Gordon.

### 2018-06-17 express2018/context.tex
Fixed all spelling issues; adjusted page breaking; deleted some really unrelated references.
text

\subsection{Context, guardness and (pre)congruence}

@@

We have chosen to use $\lambda$-expressions (typed
$CCS\rightarrow CCS$) to represent  (multi-hole) 
contexts. The definition is inductive:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{t}) \HOLSymConst{\HOLTokenConj{}} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{p}. \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{p})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{a} \HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{a}\HOLSymConst{..}\HOLBoundVar{e} \HOLBoundVar{t})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{e\sb{\mathrm{2}}}. \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{+} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{e\sb{\mathrm{2}}}. \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{L} \HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLSymConst{\ensuremath{\nu}} \HOLBoundVar{L} (\HOLBoundVar{e} \HOLBoundVar{t}))) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{rf} \HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{relab} (\HOLBoundVar{e} \HOLBoundVar{t}) \HOLBoundVar{rf})\hfill{[CONTEXT_rules]}
\end{alltt}

Under above definition, we can formally define the concept of
``precongruence'' and ``congruence'' in the following ways:
\begin{alltt}
\HOLConst{precongruence} \HOLFreeVar{R} \HOLSymConst{\HOLTokenEquiv{}}
\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{y} \HOLBoundVar{ctx}. \HOLConst{CONTEXT} \HOLBoundVar{ctx} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{y} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} (\HOLBoundVar{ctx} \HOLBoundVar{x}) (\HOLBoundVar{ctx} \HOLBoundVar{y}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{PreOrder} \HOLFreeVar{R}
\HOLConst{congruence} \HOLFreeVar{R} \HOLSymConst{\HOLTokenEquiv{}}
\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{y} \HOLBoundVar{ctx}.
    \HOLConst{CONTEXT} \HOLBoundVar{ctx} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{y} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} (\HOLBoundVar{ctx} \HOLBoundVar{x}) (\HOLBoundVar{ctx} \HOLBoundVar{y}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{equivalence} \HOLFreeVar{R}
\end{alltt}

@@

In the same manner, we can also define all
variants of above concepts without \emph{direct} sums (\texttt{GCONTEXT}, \texttt{WGS} and
\texttt{GSEQ}). A large amount of lemmas must be proved for their relationships and properties of compositions. These proofs all have long (and trivial) steps due to multiple levels of
inductions on their structures.%  For example, one such

@@

% Once the representation issue of CCS equations is resolved, the actual
% proofs of Milner's unique solution of equations theorems is not very
% interesting from the view of theorem proving, although it's a 
% precise proof engineering work producing quite long proofs.
% Since we have chosen to use  context to represent
% single-variable equations, an equation like $P \sim E\{P/X\}$ now
% becomes $P \sim E[P]$, in which $E$ is a (multi-hole) 
% context, and there's no need to say it ``contains at most the variable
% $X$'' any more, as equation variable doesn't appear in $E$ at
% all. Under these simplifications, the formal proof of Milner's unique
% solution of equations theorem for $\sim$ is a literal mapping for informal
% proofs based on bisimulation upto $\sim$, induction and case analysis
% of weakly guarded contexts. Below is the formal version of Lemma 4.13
% and Proposition 4.14 in Milner's book:

>>>>

\subsection{Context, guardedness and (pre)congruence}

@@

We have chosen to use $\lambda$-expressions (typed
$CCS\rightarrow CCS$) to represent  (multi-hole) 
contexts. The definition is inductive:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{t}) \HOLSymConst{\HOLTokenConj{}} (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{p}. \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{p})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{a} \HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{a}\HOLSymConst{..}\HOLBoundVar{e} \HOLBoundVar{t})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{e\sb{\mathrm{2}}}. \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{+} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{e\sb{\mathrm{2}}}. \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{CONTEXT} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{e\sb{\mathrm{1}}} \HOLBoundVar{t} \HOLSymConst{\ensuremath{\parallel}} \HOLBoundVar{e\sb{\mathrm{2}}} \HOLBoundVar{t})) \HOLSymConst{\HOLTokenConj{}}
   (\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{L} \HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLSymConst{\ensuremath{\nu}} \HOLBoundVar{L} (\HOLBoundVar{e} \HOLBoundVar{t}))) \HOLSymConst{\HOLTokenConj{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{rf} \HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenImp{}} \HOLConst{CONTEXT} (\HOLTokenLambda{}\HOLBoundVar{t}. \HOLConst{relab} (\HOLBoundVar{e} \HOLBoundVar{t}) \HOLBoundVar{rf})\hfill{[CONTEXT_rules]}
\end{alltt}
We can then define
`precongruence' and `congruence' thus:
\begin{alltt}
\HOLConst{precongruence} \HOLFreeVar{R} \HOLSymConst{\HOLTokenEquiv{}}
\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{y} \HOLBoundVar{ctx}. \HOLConst{CONTEXT} \HOLBoundVar{ctx} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{y} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} (\HOLBoundVar{ctx} \HOLBoundVar{x}) (\HOLBoundVar{ctx} \HOLBoundVar{y}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{PreOrder} \HOLFreeVar{R}
\HOLConst{congruence} \HOLFreeVar{R} \HOLSymConst{\HOLTokenEquiv{}}
\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{y} \HOLBoundVar{ctx}.
    \HOLConst{CONTEXT} \HOLBoundVar{ctx} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{y} \HOLSymConst{\HOLTokenImp{}} \HOLFreeVar{R} (\HOLBoundVar{ctx} \HOLBoundVar{x}) (\HOLBoundVar{ctx} \HOLBoundVar{y}) \HOLSymConst{\HOLTokenConj{}} \HOLConst{equivalence} \HOLFreeVar{R}
\end{alltt}

@@

In the same manner, we can also define all
variants of above concepts with only weakly-guarded sums (\texttt{GCONTEXT}, \texttt{WGS} and
\texttt{GSEQ}). Several lemmas have then to be proved for ther
relationships among these kinds of contexts and properties about their
compositions. These proofs are usually tedious but  long,  due to multiple levels of
inductions on the structure of the contexts.%  For example, one such

@@

% Once the representation issue of CCS equations is resolved, the actual
% proofs of Milner's unique solution of equations theorems is not very
% interesting from the view of theorem proving, although it's a 
% precise proof engineering work producing quite long proofs.
% Since we have chosen to use  context to represent
% single-variable equations, an equation like $P \sim E\{P/X\}$ now
% becomes $P \sim E[P]$, in which $E$ is a (multi-hole) 
% context, and there's no need to say it ``contains at most the variable
% $X$'' any more, as equation variable doesn't appear in $E$ at
% all. Under these simplifications, the formal proof of Milner's unique
% solution of equations theorem for $\sim$ is a literal mapping for informal
% proofs based on bisimulation up to $\sim$, induction and case analysis
% of weakly guarded contexts. Below is the formal version of Lemma 4.13
% and Proposition 4.14 in Milner's book:

### 2018-06-17 express2018/contraction.tex
Fixed all spelling issues; adjusted page breaking; deleted some really unrelated references.
text

  Intuitively,
  the bisimilarity contraction 
$\mcontrBIS$ 
 is  a preorder in which 
$P \mcontrBIS Q  $  holds  if $P \wb Q$ and, in addition, 
$Q$ has the \emph{possibility} of being at least as efficient as $P$ (as far as
$\tau$-actions performed). 
Process $Q$, however, may be nondeterministic and may have other ways
of doing the same work, and these could be  slow (i.e., involving
more $\tau$-steps than those performed by $P$).
Thus, in contrast with expansion,  we cannot really say that `$Q$ is more efficient than
$P$'.

@@

Bisimilarity  contraction is coarser than 
 the \emph{expansion  relation} 
$\expa$ \cite{sangiorgi2015equations, arun1992efficiency}. This 
is a
preorder widely used in proof techniques for bisimilarity and that 
intuitively refines bisimilarity by 
 formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the  possibility of
being at least as efficient as $P$.

>>>>

  Intuitively,
  the bisimilarity contraction 
$\mcontrBIS$ 
 is  a preorder in which 
$P \mcontrBIS Q  $  holds  if $P \wb Q$ and, in addition, 
$Q$ has the \emph{possibility} of being at least as efficient as $P$ (as far as
$\tau$-actions performed). 
Process $Q$, however, may be nondeterministic and may have other ways
of doing the same work, and these could be  slow (i.e., involving
more $\tau$-steps than those performed by $P$).
% Thus, in contrast with expansion,  we cannot really say that `$Q$ is more efficient than
% $P$'.

@@

Bisimilarity  contraction is coarser than 
 the \emph{expansion  relation} 
$\expa$ \cite{sangiorgi2015equations} %, arun1992efficiency}.
This is a
preorder widely used in proof techniques for bisimilarity and that 
intuitively refines bisimilarity by 
 formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the  possibility of
being at least as efficient as $P$.

### 2018-06-17 express2018/equa.tex
text

\begin{definition}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ possibly containing  such variables. 
Then 
$$\{  X_i = E_i\}_{i\in I}$$
is 
  a \emph{system of equations}. (There is one equation for each variable $X_i$.)
\end{definition}

@@

For instance, the solution of the equation 
$ X = a. X$ 
is  the process
$R \DSdefi \recu A { a. A}$, and   for any other solution $P$ we have $P \wb R$.
In contrast, the equation 
 $X = a|  X$ has solutions that may be quite different, namely any process capable of
continuously  performing $a$ action (but that could behave arbitrarily  on other actions).

>>>>

\begin{definition}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ possibly containing  such variables. 
Then 
$\{  X_i = E_i\}_{i\in I}$
is 
  a \emph{system of equations}. (There is one equation for each variable $X_i$.)
\end{definition}

@@

For instance, the solution of the equation 
$ X = a. X$ 
is  the process
$R \DSdefi \recu A { a. A}$, and   for any other solution $P$ we have $P \wb R$.
In contrast, the equation 
 $X = a|  X$ has solutions that may be quite different, namely any process capable of
continuously  performing $a$ actions (but that could behave arbitrarily  on other actions).

### 2018-06-17 express2018/expansion.tex
Make some more spaces
text

\subsection{Bisimilarity, rooted bisimilarity,  and up-to techniques}
\label{ss:BiEx}

@@

We sometimes call bisimilarity as ``\emph{weak}'', to
distinguish it from \emph{strong bisimilarity}, $\sim$,
obtained by replacing in the above definition   the weak answer $
Q\Arcap\mu Q'$ with the strong  $Q \arr \mu Q'$.

Bisimilarity is not preserved by the sum operator. 
For this, Milner introduces \emph{observational congruence}, also called \emph{rooted
  bisimilarity}.
\begin{definition}[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written $P
\rapprox Q$, if we have:
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
\item the converse of (1) on the actions from $Q$.
\end{enumerate}
\end{definition}
% Besides reducing the rooted bisimiarity of two processes to
% the bisimilarities of their first-step transition ends, this definition also brings a proof technique for proving the
% rooted bisimiarity by constructing a bisimulation:
% \begin{lemma}{(Rooted bisimilarity by constructing a bisimulation)}
% \label{l:obsCongrByWeakBisim}
% Given a (weak) bisimulation $\RR$, if two processes $P$ and $Q$
% satisfies the following properties:
% \begin{enumerate}
%  \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
%    \Arr\mu Q'$ and $P' \RR Q'$;
% \item the converse of (1) on the actions from $Q$.
% \end{enumerate}
% then $P$ and $Q$ are rooted bisimilar, i.e., $P \approx^c Q$.
% \end{lemma}

\begin{theorem}
\label{t:rapproxCongruence}
Relation $\rapprox$ is a congruence in CCS.  Indeed it is the
largest congruence contained in $\approx$.
\end{theorem}

>>>>

\subsection{Bisimilarity and  rooted bisimilarity}
\label{ss:BiEx}

@@

We sometimes call bisimilarity `\emph{weak}', to
distinguish it from \emph{strong bisimilarity}, $\sim$,
obtained by replacing in the above definition   the weak answer $
Q\Arcap\mu Q'$ with the strong  $Q \arr \mu Q'$.
Bisimilarity is not preserved by the sum operator. 
For this, Milner introduces \emph{observational congruence}, also called \emph{rooted
  bisimilarity}.
\begin{definition}[rooted bisimilarity]
\label{d:rootedBisimilarity}
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written $P
\rapprox Q$, if we have:
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
\item the converse of (1) on the actions from $Q$.
\end{enumerate}
\end{definition}
% Besides reducing the rooted bisimiarity of two processes to
% the bisimilarities of their first-step transition ends, this definition also brings a proof technique for proving the
% rooted bisimiarity by constructing a bisimulation:
% \begin{lemma}{(Rooted bisimilarity by constructing a bisimulation)}
% \label{l:obsCongrByWeakBisim}
% Given a (weak) bisimulation $\RR$, if two processes $P$ and $Q$
% satisfies the following properties:
% \begin{enumerate}
%  \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
%    \Arr\mu Q'$ and $P' \RR Q'$;
% \item the converse of (1) on the actions from $Q$.
% \end{enumerate}
% then $P$ and $Q$ are rooted bisimilar, i.e., $P \approx^c Q$.
% \end{lemma}

\begin{theorem}
\label{t:rapproxCongruence}
$\rapprox$ is a congruence in CCS. Indeed, it is the
largest congruence contained in $\approx$.
\end{theorem}

### 2018-06-17 express2018/formal.tex
text

\section{Formalisation}
\label{s:for}
We highlight here a formalisation of CCS
in the HOL theorem
prover (HOL4) \cite{slind2008brief},
including the new concepts and theorems proposed in the first half of
this paper.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
The whole formalisation can be found 
in \cite{Tian:2017wrba}  or in 
 in the official example folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}. The
work consists of about 20,000 lines of proof scripts in Standard ML
(and that follows the pioneering work by Nesi \cite{Nesi:1992ve}, in
1992-1995).

Higher Order Logic (or ``HOL Logic'') \cite{hollogic} stands for simple-typed $\lambda$-calculus plus Hibert
choice operator, axiom of infinity, and rank-1 polymorphism (type
variables). HOL4 implements the original HOL logics, 
in contrast with 
 other theorem provers of the HOL family (e.g. Isabelle/HOL) that have
made extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed the HOL Logic has considerable simpler logic
foundations than most other theorem provers. %, e.g. Coq. 
As a result,
formal theories in HOL can be easily migrated (sometimes even
automatically) into other theorem provers.

HOL4 is written in Standard ML (ML for abbreviation), which plays three roles here:
\begin{enumerate}
\item It serves as the underlying implementation language for the core HOL engine;
\item it is used to implement tactics (and tacticals) for writing proofs;
\item it is used as the command language of the HOL interactive shell.
\end{enumerate}
Besides, HOL4 users usually write complex automatic
verification tools (in the same language) by calling HOL's interactive theorem proving facilities.
\iflong
The HOL logic language, on the other
 side, is more close to Classic ML, in which the early HOL systems were built.
\DS{removed the sentence because  it did not
  seem important} 
\fi

@@

In this
formalisation we consider only single equations/contractions. 
This setup considerably simplified the required proof steps while still aligned with
their informal proofs.
%  (on 
% the extension  to multiple equations/contractions
% see also  Section~\ref{s:concl}).

>>>>

\section{Formalisation}
\label{s:for}
We highlight here a formalisation of CCS
in the HOL theorem
prover (HOL4) \cite{slind2008brief},
including the new concepts and theorems proposed in the first half of
this paper.
%  The main purpose is to convince the readers that, there's no flaw
%  in the informal proofs. 
The whole formalisation can be found 
in \cite{Tian:2017wrba}  or in 
 in the official example folder of HOL4 source
code\footnote{\url{https://github.com/HOL-Theorem-Prover/HOL}}. The
work consists of about 20,000 lines of proof scripts in Standard ML;
it develops
some pioneering work by Nesi \cite{Nesi:1992ve}, in
1992-1995.

Higher Order Logic (or ``HOL Logic'') \cite{hollogic} stands for simple-typed $\lambda$-calculus plus Hilbert
choice operator, axiom of infinity, and rank-1 polymorphism (type
variables). HOL4 implements the original HOL logics, 
in contrast with 
 other theorem provers of the HOL family (e.g. Isabelle/HOL) that have
made extensions.
%  (they made the formal language more powerful,
% but they also bring the possibilities that the entire logic becomes
% inconsistent). 
Indeed the HOL Logic has considerable simpler logic
foundations than most other theorem provers. %, e.g. Coq. 
As a result,
formal theories in HOL can be easily migrated (sometimes even
automatically) into other theorem provers.

HOL4 is written in Standard ML (ML for abbreviation), which plays three roles here:
\begin{enumerate}
\item It serves as the underlying implementation language for the core HOL engine;
\item it is used to implement tactics (and tacticals) for writing proofs;
\item it is used as the command language of the HOL interactive shell.
\end{enumerate}
Further, in the same language
 HOL4 users can write complex automatic
verification tools  by calling HOL's interactive theorem proving facilities.
\iflong
The HOL logic language, on the other
 side, is more close to Classic ML, in which the early HOL systems were built.
\DS{removed the sentence because  it did not
  seem important} 
\fi

@@

In this
formalisation we consider only single equations/contractions. 
This  %considerably
 simplifies the required proof steps in HOL   while enhancing the
 reading of the scripts.
%   still aligned with
% their informal proofs.
%  (on 
% the extension  to multiple equations/contractions
% see also  Section~\ref{s:concl}).
